Solution of Differential Algebraic Equations with Local Constraints on the Latent Space of an Autoencoder

### 2019-06-22 paper/paper.tex
edited the paper

@@

\usepackage{natbib}

@@

% set default figure placement to htbp
\makeatletter
\def\fps@figure{htbp}
\makeatother

\title{Solution of Differential Algebraic Equations with Manifold Constraints
on the Latent Space of an Autoencoder}
\author{Alejandro Francisco Queiruga}
\date{\today}

@@

Representing complex physical systems with multiple speicies, phases,
chemical reactions, and heterogenous scale effects poses an
intractability problem to modeling. How do we analyze complex
experimental data and \emph{ab initio} simulations and derive new
macroscale relations to describe the systems? How do we then implement
these relations in the rest of our analyses and production-scale
computer simulations?

Accurately simulating multispecies reaction and transport with
undergoing phase transitions is a critical aspect to a variety of
problems.
This work is motivated by advancing simulation techniques for subsurface energy systems such as shale
and oceanic natural gas reservoirs in shales or engineered geothermal
systems.

@@

% {\bf ML LIT REVIEW HERE} % TODO
\cite{wu_physics-informed_2018}
\cite{xie_tempogan:_2018}

@@

% \begin{figure}
%    CSTR
%   \caption{ \label{fig:cstr} Illustration of the Continuous stirred-tank
%   reactor. The material in the reactor is exposed to a reservoir with pressure
%   $p_\inf$ and temperature $T_\inf$ via a conduit with permeability
%   and thermal conductivities $k_p$ and $k_T$, direct forced mass
%   injection/withdrawl $r$, or heating and cooling $s$.}
% \end{figure}

For this preliminary work, we consider only a single-beaker problem
without flow, a continuous stirred-tank reactor (CSTR), as illustrated
in Figure \ref{fig:cstr}. Let \(\rho\) be the material density, \(T\) be the
temperature, \(p\) be the material pressure, \(h\) be the material
enthalpy, and \(u\) be the internal energy. There are two equations for
balance of mass and energy.
\begin{align}
\partial_t \rho & = \mathbf{k}_p(p_\infty - p) + r\\
\partial_t \rho u & = \mathbf{k}_T(T_\infty-T) + s 
\end{align}
The beaker is exposed to a reservoir with \(p_\infty\) and \(T_\infty\)
via a mixed boundary condition straw with transport and conduction
coefficients \(\mathbf{k}_p\), \(\mathbf{k}_T\). Mass and heating
sources and sinks are the terms \(r\) and \(s\).\footnote{We express the
  material by enthalpy instead of internal energy; they are related by
  the relation \(\rho u = \rho h - p\). Future work may work directly
  with internal energy.}
We are neglecting upstreaming in the connection between the reservoir
and the container.

@@

We can also assemble the surface of possible $T,p,\rho$ coordinates,
shown in Figure \ref{fig:pTrho}.
\begin{figure}
\centering
\includegraphics[width=4in]{/Users/afq/Documents/Dropbox/ML/primaryautoencoder/figures/phase_diagram.png}
\caption{\label{fig:phaseboundaries}Phase boundaries for water on the $T-p$ diagram. The standard
interpretation is that the properties of the material change
drastically when crossing this boundary. Note how it is possible to form a path from a
definitively gas state to a definitively liquid state without crossing
a phase boundary. The supercritical regime has no clear boundaries,
but requires a seperate curve fit in practice due to numerical challenges.}
\end{figure}

\begin{figure}
\centering
\includegraphics[width=4in]{/Users/afq/Documents/Dropbox/ML/primaryautoencoder/figures/water_eos.png}
\caption{\label{fig:pTrho} The $\log{p}$, $T$, $\rho$ equation of
  state surface for water in units of $\log{\text{Pa}}$, K, and \si{}{kg/m$^2$} This plot spans four phases, three
  equilibria, and the triple point, which are each given their own color.
  There is a fourth dimension to the surface of enthalpy $h$ that exhibits
  similar discontinuities; the $\log{p}$, $T$, $h$ surface is plotted
  in the Appendix. (Every point on this plot is a point in the
  generated dataset used for the optimization)}
\end{figure}

{\bf Remark:}There are even more known phases of crystal ice which for water
For water, the supercritical phase and exotic ices only occur in
laboratory conditions or naturally near the
crust \cite{}. For our case, we only handle the Ice I and cutoff the
regime at \si{10^8}{Pa}. For other materials these exotic phases can be
applicable. For example, in geoengineered systems using $CO_2$
such as geothermal plants are carbon sequestration facilities, the
working fluid is usually in the supercritical regime at depth. cite
curt. cite hydrates.
When binary mixtures are involved, the number of possible phases and
equilibria increases again.

@@

@@

The typical methodology is to use the empirical relations for density
and enthalpy \(h\) as a function of pressure and temperature and solve
the DAE for \(p\) and \(T\) implicitly:
\begin{align}
\text{Solve for}\, p(t)\, \text{and}\, T(t)\, \text{such that:}\\
\partial_t \rho(p,T) & = \nabla \cdot \mathbf{k}\nabla p + r\\
\partial_t \rho u(p,T) & = \nabla \cdot \mathbf{k'}\nabla T + s
\end{align}
The complicatation is that the functions \(\rho(p,T)\) and \(u(p,T)\)
are not well defined functions due to the presence of phase changes that
yield sharp discontinuities in \(p,T\), as shown in Figure 1 for
water.

@@

  This requires a statemachine. For the complete states of water, it is shown in Figure \ref{fig:statemachine}.
  \begin{figure}
    \includegraphics[width=4in]{/Users/afq/Documents/Dropbox/ML/primaryautoencoder/slides/phase_4.png}
    \caption{\label{fig:statemachine}States and possible transitions for water.}
\end{figure}
Formulate balance equations as
\begin{equation}
R(X ; phase) = 0
\end{equation}
Solve the differentiable part:
\[\frac{\partial R}{\partial X}\Delta X^{k+1/2} = -R(X^k ; phase^k)\]
and iterate on the nondifferential part:
\[phase^{k+1},X^{k+1} = statemachine(X^{k+1/2} ; phase^k)\]
The phase switching logic has the form of a switch case:

@@

Soltutions are shown in Figure \ref{fig:pendulum}.
\begin{figure}
\centering
\includegraphics[width=5in]{/Users/afq/Documents/Dropbox/ML/primaryautoencoder/figures/pendulum_q.png}
\caption{\label{fig:pendulum}Solution of the pendulum on an arbitrary latent space.}
\end{figure}

@@

The architecture is illustrated in Figure \ref{fig:classifyingnetwork}.
It has the following form in index notation:
\begin{align}
  q_i &= W^{enc}_{ij} x^n_j + b^{enc}_i\\
  f_{ij} &= W^{dec}_{ijk} q^m_k + b^{dec}_{ij} \\
  h_i &= \mu\left(W^{bound}_{ik}q^m_k+b^{bound}_{i}\right) \\
  p_i &= \sigma\left(W^{phase}_{ij}h_j+b^{phase}_i\right) \\
  \bar{x}_i&= f_{ij} p_j
\end{align}
where $x$ is the original prediction, $x^n$ denotes a polynomial basis
set on $x$ (which increases the number of channels), $q$ is the latent
space, $f$ is a set of polynomial curves, $h$ is a hidden layer
denoting sides of boundaries, $p$ is a classification layer depending
on $h$, and $\bar{x}$ is the predicted feature. The activation $\mu$
is chosen as a standard neuron, e.g. $\tanh$ or $\text{sigmoid}$. The
activation $\sigma$ is a probability activation, which is chosen to be
a softmax,
\begin{equation}
\sigma_i(x;\beta) = \frac{e^{\beta x_i}}{\sum_j e^{\beta x_j}}
\end{equation}
or, in the limit, the hard argmax:
\begin{equation}
\sigma_i \rightarrow \text{onehot}\arg\max z as  \beta\rightarrow\infty
\end{equation}

\begin{figure}
\centering
\includegraphics{/Users/afq/Documents/Dropbox/ML/primaryautoencoder/slides/classifier_network.png}
\caption{\label{fig:classifyingnetwork}The architecture of the classifier network designed for
  unsupervised labeling of phases.}
\end{figure}

\begin{figure}
  \centering
  \includegraphics[width=3in]{/Users/afq/Documents/Dropbox/ML/primaryautoencoder/slides/classifier_good.jpg}
  \caption{\label{fig:classified}Result of the unsupervised phase classifier. Compare this plot to the original labeled dataset Figure \ref{fig:pTrho}.}
\end{figure}
Because the identity of the phases,XXXXXXXX
For comparison purposes, we group together liquid-gas-supercritical as one contiguous label.

@@

To develop the rest of the discretization, we introduce the compact form:
\begin{align}
\frac{\mathrm{d}}{\mathrm{d}t} m(T,p,\rho,h) &= r(T,p,\rho,h)\\
eos(T,p,\rho,h) &= 0
\end{align}
where $m=\{\rho,\rho h -p\}$ and $r$ is the right hand side, with no good representation of \(eos\) everywhere.

The decoder part of the network is inserted into the balance law
\begin{equation}
\frac{\mathrm{d}}{\mathrm{d}t} m(D(q)) = r(D(q))
\end{equation}
We can parameterize this equation with any time stepping scheme, .e.g. a
Runge-Kutta scheme. For subsurface flow, backward Euler is popular due
to its simplicity and L-stability, which we follow since the topic of
this paper is not exploring new time stepping schemes. This yields the
discretizated algebraic equation for \(q_i\) and \(t+\Delta t\) given the
value \(q^0\) at \(t\),

@@

@@

\hypertarget{header-n3356}{%
\section{Implementation}\label{header-n3356}}

@@

The codebase is released open source at
\url{https://github.com/afqueiruga/latentsim}. Helper methods and
custom optimizers that the
author reuses in various TensorFlow-based projects are factored out in
the dully-named \url{https://github.com/afqueiruga/afqstensorflowutils}.
The implementation of the empirical equations of state for water are
factored out at
\href{https://github.com/afqueiruga/equations_of_state}{https://github.com/afqueiruga/equations\_of\_state}.

@@

In TOUGH, for one instance, all of the calculations are performed in SI units because it is simply too much labor to normalize everything! Adverse effects to iterative algorithms and floating point error propogation occur, especially because numerical differentiation is used.
The learned primary variables are normalized to a unit range around zero, and the scaling matrix is explicitly in one step. The scaling matrix could be a way to systematically non-dimensionalize the entire problem.

@@

Techniquessuch as L1 regularization would be needed to search for low
cost models that also satisfy our desire for Occam's razor
simplifications. (The IAPWS relations used a prelimary form of this
type of simplification, but are not optimal in the author's opinion.)

>>>>

\usepackage{siunitx}

@@

\usepackage[numbers]{natbib}

@@

% \makeatletter
% \def\fps@figure{tbph}
% \makeatother

\title{Solution of Differential Algebraic Equations with Local Constraints
on the Latent Space of an Autoencoder}
\author{Alejandro Francisco Queiruga}
\date{\today}

@@

Representing complex physical systems with multiple species, phases,
chemical reactions, and heterogenous scale effects poses an
intractability problem to modeling. How do we analyze complex
experimental data and \emph{ab initio} simulations and derive new
macroscale relations to describe the systems? How do we then implement
these relations in the rest of our analyses and production-scale
computer simulations
Accurately simulating multispecies reaction and transport with
undergoing phase transitions is a critical aspect to a variety of
problems.
This work is motivated by advancing simulation techniques for subsurface energy systems such as shale
and oceanic natural gas reservoirs in shales or engineered geothermal
systems.

Our fundamental goal is to develop the tools
for unsupervised representation learning in way where the
representations have physical and useful meaning.
How can we find a latent space that is good for our tasks?
How can we find regressions in a deep learning context of high enough
accuracy with needed qualities?
How do we include these types of machine-generated models into a
scientific and engineering calculation without sacrificing accuracy?
Can this improve our workflow?
Can we then find paths in the learned space that follow physical laws?

@@

% {\bf ML LIT REVIEW HERE} % TODO
Representation learning methods such as StyleGAN and G|FLOW have shown
the ability to find meaningful dimensions in the latent space,
e.g. the popular ability to transform age, race, and gender of human
faces in a smooth way.
But, can this directions be used meaningfully?

\cite{wu_physics-informed_2018}
\cite{xie_tempogan:_2018}

@@

For this preliminary work, we consider only a single-beaker problem
without flow, the idealization of the continuous stirred-tank reactor (CSTR), as illustrated
in Figure \ref{fig:cstr}. Let \(\rho\) be the material density, \(T\) be the
temperature, \(p\) be the material pressure, \(h\) be the material
enthalpy, and \(u\) be the internal energy. There are two equations for
balance of mass and energy.
\begin{align}
\partial_t \rho & = \mathbf{k}_p(p_\infty - p) + r\\
\partial_t \rho u & = \mathbf{k}_T(T_\infty-T) + s 
\end{align}
The beaker is exposed to a reservoir with \(p_\infty\) and \(T_\infty\)
via a mixed boundary condition straw with transport and conduction
coefficients \(\mathbf{k}_p\), \(\mathbf{k}_T\). Mass and heating
sources and sinks are the terms \(r\) and \(s\).\footnote{We express the
  material by enthalpy instead of internal energy; they are related by
  the relation \(\rho u = \rho h - p\). Future work may work directly
  with internal energy.}
We are neglecting upstreaming in the connection between the reservoir
and the container.

\begin{figure}
  \centering
  \includegraphics[width=3in]{../slides/cstr.png}
  \caption{ \label{fig:cstr} Illustration of the Continuous
    Stirred-Tank    
  Reactor. The material in the reactor is exposed to a reservoir with
  pressure
  $p_\infty$ and temperature $T_\infty$ via a conduit with permeability
  and thermal conductivities $k_p$ and $k_T$, direct forced mass
  injection or withdrawal $r$ via a pump, or heating and cooling $s$
  via a heater/heat-exchanger. Stirring and insulation keeps the
  material at a uniform pressure and temperature. Gravity is
  neglected (though is usually the motivation for putting a feeding inlet at
the top and product outlet at the bottom.)}
\end{figure}

@@

@@

\subsection{Forms of Equations of State}

This work focuses on the EOS of pure water. The most accurate
formulations were developed by the International Association for the
Properties of Water Steam (IAPWS) using a variety of experimental
sources for multiple phases \cite{wagner_iapws_2000,
  water_revised_2007, water_revised_2009}. We utilize the 1997
formulation for water gas, liquid, and supercritical phases, and the
2006 formulation for the Ice Ih phase to build the dataset.\footnote{Our
implementation is publicly available at GITHUB:EOS.} (Their
painstaking work in determining the forms of decent fits is a great
motivation for a 21st century approach.)
We can also assemble the surface of possible $T,p,\rho$ coordinates,
shown in Figure \ref{fig:pTrho}.
\begin{figure}
\centering
\includegraphics[width=4in]{../figures/phase_diagram.png}
\caption{\label{fig:phaseboundaries}Phase boundaries for water on the $T-p$ diagram. The standard
interpretation is that the properties of the material change
drastically when crossing this boundary. Note how it is possible to form a path from a
definitively gas state to a definitively liquid state without crossing
a phase boundary. The supercritical regime has no clear boundaries,
but requires a separate curve fit in practice due to numerical challenges.}
\end{figure}

\begin{figure}
\centering
\includegraphics[width=4in]{../figures/water_eos.png}
\caption{\label{fig:pTrho} The $\log{p}$, $T$, $\rho$ equation of
  state surface for water in units of $\log{\text{Pa}}$, K, and \si{}{kg/m$^2$} This plot spans four phases, three
  equilibria, and the triple point, which are each given their own color.
  There is a fourth dimension to the surface of enthalpy $h$ that exhibits
  similar discontinuities; the $\log{p}$, $T$, $h$ surface is plotted
  in the Appendix. (Every point on this plot is a point in the
  generated dataset used for the optimization)}
\end{figure}

{\bf Remark:} There are even more known phases of crystal ice which for water
For water, the supercritical phase and exotic ices only occur in
laboratory conditions or naturally near the
crust \cite{}. For our case, we only handle the Ice I and cutoff the
regime at \si{10^8}{Pa}. For other materials these exotic phases can be
applicable. Methane $CH_4$ is actually well beyond its supercritical
point at T=\si{-82.59}{C}\cite{the_engineering_toolbox_methane_2008}, so only an appropriate gas law can do the trick, until
the bi-species hydrate phase comes into play.
The liquid-gas-equilibrium-supercritical state transition comes into
play in, for example, in geoengineered systems using $CO_2$
such as geothermal plants are carbon sequestration facilities, the
working fluid is often in the supercritical regime at depth, and the
phase boundary. cite
curt. cite hydrates.
When binary mixtures are involved, the number of possible phases and
equilibria increases again.

{\bf Remark:} Many popular forms of EOS do not even have closed form
expressions. The family of cubic equations of state for nonideal
gases, such as a Peng-Robinsion equation \cite{peng_new_1976}, are
expressed as functions for $p$ of $V$ (a proxy for $\rho$) and
$T$. Since $\rho$ is rarely a primary varibale, these expressions must be inverted for
$\rho$ given $p$ by solving the cubic equation. This can be performed
analytically, as is done in TOUGH+, but often polynomial roots can
be best solved numerically. When the equations become more
complicated, such as when there are gas mixtures, the equations of
state must be solved numerically. TOUGH+ employs a Richardson
iteration \cite{moridis_users_2014-2}.
Note how this is a numerical
iteration {\em inside} of another numerical iteration that is trying
to differentiation this equation state. Approximations
abound--the mathematical theory is not straight foward. Passing
automatic differentiation objects through these computations is
difficult in practice. Equations are
often solved faster and more accurately by numerical
approximations--why not completely rewrite the equations?

{\bf Remark:} The relations inside of EOSes are often artificially
smoothed. For example, the hydrate equilibrium surface has a kink at
the ice-liquid phase boundary in water. This causes problems
numerically when a material state is moving along this equilibrium, so
it is artificially smoothed using splines or the hyperbolic tangent to
improve convergence in TOUGH+Hydrate \cite{moridis_users_2014}. Approximations abound.

@@

The typical methodology is to use the empirical relations for density
and enthalpy \(h\) as a function of pressure and temperature and solve
the DAE for \(p\) and \(T\) implicitly:
\begin{align}
\text{Solve for}\, p(t)\, \text{and}\, T(t)\, \text{such that:}\\
\partial_t \rho(p,T) & = \nabla \cdot \mathbf{k}\nabla p + r\\
\partial_t \rho u(p,T) & = \nabla \cdot \mathbf{k'}\nabla T + s
\end{align}
The complication is that the functions \(\rho(p,T)\) and \(u(p,T)\)
are not well defined functions due to the presence of phase changes that
yield sharp discontinuities in \(p,T\), as shown in Figure 1 for
water.

@@

  This requires a statemachine. For the complete states of water, it is shown in Figure \ref{fig:statemachine}.
  \begin{figure}
    \centering
    \includegraphics[width=4in]{../slides/phase_4.png}
    \caption{\label{fig:statemachine}States and possible transitions for water.}
\end{figure}
Formulate balance equations as
\begin{equation}
R(X ; phase) = 0
\end{equation}
Solve the differentiable part:
\[\frac{\partial R}{\partial X}\Delta X^{k+1/2} = -R(X^k ; phase^k)\]
and iterate on the nondifferential part:
\[phase^{k+1},X^{k+1} = statemachine(X^{k+1/2} ; phase^k)\]
The phase switching logic has the form of a switch case:

@@

Soltutions are shown in Figure \ref{fig:pendulum}.
\begin{figure}
\centering
\includegraphics[width=5in]{../figures/pendulum_q.png}
\caption{\label{fig:pendulum}Solution of the pendulum on an arbitrary latent space.}
\end{figure}

@@

\begin{figure}
\centering
\includegraphics{../slides/classifier_network.png}
\caption{\label{fig:classifyingnetwork}The architecture of the classifier network designed for
  unsupervised labeling of phases.}
\end{figure}

The architecture is illustrated in Figure \ref{fig:classifyingnetwork}.
It has the following form in index notation:
\begin{align}
  q_i &= W^{enc}_{ij} x^n_j + b^{enc}_i\\
  f_{ij} &= W^{dec}_{ijk} q^m_k + b^{dec}_{ij} \\
  h_i &= \mu\left(W^{bound}_{ik}q^m_k+b^{bound}_{i}\right) \\
  p_i &= \sigma\left(W^{phase}_{ij}h_j+b^{phase}_i\right) \\
  \bar{x}_i&= f_{ij} p_j
\end{align}
where $x$ is the original prediction, $x^n$ denotes a polynomial basis
set on $x$ (which increases the number of channels), $q$ is the latent
space, $f$ is a set of polynomial curves, $h$ is a hidden layer
denoting sides of boundaries, $p$ is a classification layer depending
on $h$, and $\bar{x}$ is the predicted feature. The activation $\mu$
is chosen as a standard neuron, e.g. $\tanh$ or $\text{sigmoid}$. The
activation $\sigma$ is a probability activation, which is chosen to be
a softmax,
\begin{equation}
\sigma_i(x;\beta) = \frac{e^{\beta x_i}}{\sum_j e^{\beta x_j}}
\end{equation}
or, in the limit, the hard argmax:
\begin{equation}
\sigma_i \rightarrow \text{onehot}\arg\max z \quad as\quad  \beta\rightarrow\infty
\end{equation}

\begin{figure}
  \centering
  \includegraphics[width=3in]{../slides/classifier_good.jpg}
  \caption{\label{fig:classified}Result of the unsupervised phase classifier. Compare this plot to the original labeled dataset Figure \ref{fig:pTrho}.}
\end{figure}
Because the identity of the phases,XXXXXXXX
For comparison purposes, we group together liquid-gas-supercritical as one contiguous label.

@@

To develop the rest of the discretization, we introduce the compact form:
\begin{align}
\partial_t m(T,p,\rho,h) &= r(T,p,\rho,h)\\
eos(T,p,\rho,h) &= 0
\end{align}
where $m=\{\rho,\rho h -p\}$ and $r$ is the right hand side, where
$eos(T,p,\rho,h)=0$ is the material constraint.

The decoder part of the network is inserted into the balance law
\begin{equation}
\partial_t (D(q)) = r(D(q))
\end{equation}
We can parameterize this equation with any time stepping scheme, .e.g. a
Runge-Kutta scheme. For subsurface flow, backward Euler is popular due
to its simplicity and L-stability, which we follow since the topic of
this paper is not exploring new time stepping schemes. This yields the
discretized algebraic equation for \(q_i\) and \(t+\Delta t\) given the
value \(q^0\) at \(t\),

@@

\begin{figure}
  \centering
  \includegraphics[width=6in]{../slides/user_workflow.png}
  \caption{\label{fig:workflow}Work flow of the simulation programming process.}
\end{figure}
This process yields are more complicated simulation software system
than previous generations. The architecture and workflow from problem
specification to user experience is illustrated in Figure
\ref{fig:workflow}.
Previous generations of simulators were entirely composed of human
written computer code, with some inline parameters (such as those of
the IAPWS fits typed into constants in dedicated fortran files)
fits and the occasional parameter files that are fed into the system
as inputs (such as reaction rates databases.)
Now, the behavior of the software is not only defined by the human
written code, it is also defined by the model it loads, and the
dataset that produced it.

In our vision, the theoretician or experimentalist specify a dataset
defined the possible states of the material. Multiple architectures
are trained using the autoencoder unsupervised representation learning
goal. Then, the simulation is ran by loading in each of the models and tested against known problems that
span the problem space. Some of the tests are written by the original
developers, and some of them must be specified by the scientist to
pass through the newly added physical regimes.
This testing process detects models that may have faired well in the
ML process and produced low values of the loss function on testing and
validation, but cannot be actually used in the simulation.
This testing process is also used to determine the model that is
optimal in terms of compute time and satisfactorily accurate.
The ``best'' model can then be ``shipped'' (i.e., marked for
deployment in DVC) to be automatically selected by the final end user,
who only specifies the Equation of State.

What is not shown in this figure is the testing of the data, and the
testing of the simulation framework, which are needed in a
fully-fledged machine learning based software system \cite{google ML smells}.
Testing of the data is performed before the simulation
methodology. These are contained in the Equations\_Of\_State package in our case, and would
need to be performed by the theoretician and experimentalist feed-side
users.
Integration testing of the train-deploy-load-simulation framework is
performed using the linear equation of state as a trivial test dataset
with a trivial linear model.

@@

\hypertarget{header-n3356}{%
\section{Architecture}\label{header-n3356}}

The use of ML models yields a new software structure paradigm. It is now much simpler to implement new equation of state modules in
the simulation package, but managing the datasets and models adds
another layer to the software management. The high-level architecture
and low-level implementation details developed for the prototype
LatentSim are described in this section.

LatentSim is implemented entirely in Python 3 using the

The original datasets are written out as .csv files. To prep the
datasets, the numbers are scaled to a zero-centered unit-range, shuffled, and broken up into the
2/3-1/6-1/6 subdivisions for train, test, and validate subsets. The
sets are stored as simple .npy files, with an accompanying short csv file
for the scaling (which is loaded by the simulation to de-scale the
network outputs.)

The ``hub'' has a directory structure:
\begin{verbatim}
  1. training\_[EOS]
    - One folder per architecture
      - MonitoredTraningSession checkpoints
      - Tensorboard summaries
      - .csv 3d plot renderings
    - More architectures...
  2. training\_[other EOS]...
  3. test\_database
    - One sqlite file per EOS containing all tests
  4. report
    - static renderings (superseded by Dash app)
\end{verbatim}
The code has the following structure:
\begin{verbatim}
  1. sources for data
  2. prepped data\_files
  3. reference solutions for testing
  4. training code
  5. LatentSim.py
  6. testing code
  7. Dash app
\end{verbatim}
For every EOS, there are multiple generated implementations.
After using a combination of Tensorboard, [] Visdom, [] and Paraview [] during
development, a custom Dash [] web interface was ultimately developed
to asses the results. A screenshot interface is shown in Figure []. The
interface allows the developer (i.e., the author) to browse all
statistics, renderings, of the trained surfaces and their test
problems.

In the version of the software distributed to users, only one
``frozen'' and optimized model needs to be included per equation of state.

\subsection{Implementation details}

@@

The codebase is released open source at
\url{https://github.com/afqueiruga/latentsim}. Datasets and other
presentation artifacts such as videos and interactive plots are stored
in git-lfs. Helper methods and
custom optimizers that the
author reuses in various TensorFlow-based projects are factored out in
the dully-named \url{https://github.com/afqueiruga/afqstensorflowutils}.
The implementation of the empirical equations of state for water are
factored out at
\href{https://github.com/afqueiruga/equations_of_state}{https://github.com/afqueiruga/equations\_of\_state}.

\subsection{Things we'll do differently}

Adopt julia for at least the simulation

@@

In TOUGH, for one instance, all of the calculations are performed in SI units because it is simply too much labor to normalize everything! Adverse effects to iterative algorithms and floating point error propagation occur, especially because numerical differentiation is used.
The learned primary variables are normalized to a unit range around zero, and the scaling matrix is explicitly in one step. The scaling matrix could be a way to systematically non-dimensionalize the entire problem.

@@

Techniques such as L1 regularization would be needed to search for low
cost models that also satisfy our desire for Occam's razor
simplifications. (The IAPWS relations used a preliminary form of this
type of simplification, but are not optimal in the author's opinion.)
